---
title: "EDA Roy"
author: "Roy Madpis"
date: "2023-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
#library(GGally) #for the ggparcoord
#library(plotly) #for interactive plots
library(vcd) #for mosaic plots

############### Roy Madpis Personal theme setting: #################
# Using devtools and ggthemr (from github) to get a fabulous theme
library(devtools)
#devtools::install_github('cttobin/ggthemr') #need to run this line only if you don't ran it in the past
library(ggthemr)
### there are 17 built-in themes with ggthemr, all are listed in the github site: https://github.com/Mikata-Project/ggthemr#palettes
### I like the "flat" and the "flat dark" option :)
ggthemr("flat") #flat
#ggthemr_reset() #if you wish to return to the "normal" setting, run this line
########################################################################

library(dplyr)

#install.packages("remotes")
remotes::install_github("jtr13/redav")
library("redav")



```

## R Markdown


```{r}
data <- read_csv("data/Crash_Reporting_-_Incidents_Data.csv")

### Preprocess:
new_col_names <- colnames(data)
# Replace spaces with underscores
new_col_names <- gsub(" ", "_", new_col_names)
new_col_names <- gsub("-", "_", new_col_names)
colnames(data) <- new_col_names

data |> head()
```

```{r}
# Function to replace "N/A" with NA in a vector
replace_na <- function(x) {
  x[x == "N/A"] <- NA
  return(x)
}

# Apply the function to each column of the data frame
data <- as_tibble(lapply(data, replace_na))
```



### Analyzing NAs

```{r}
b <- apply(data, 2, function(c) sum(is.na(c)))
tibble_nas <- tibble(column_name = names(b), number_of_NA = b)
tibble_nas

ggplot(tibble_nas |> filter(number_of_NA >0 ))+
  geom_col(mapping = aes(y=reorder(column_name,number_of_NA), x = number_of_NA)) +
  labs(title = "Number of Nas in the data columns",
       subtitle = "Showing only columns with number of NAs > 0")

```



```{r}

paste0("Number of columns with at least 1 NA = ", tibble_nas |> filter(number_of_NA > 0) |> nrow())

paste0("Number of columns without any NAs = ", tibble_nas |> filter(number_of_NA == 0) |> nrow())

```

There are 9 columns with "too many" NAs (more than 45% of the rows)
```{r}
columns_with_too_many_nas <- tibble_nas|> filter(number_of_NA>0) |> arrange(number_of_NA) |> filter(number_of_NA > 45000)
columns_with_too_many_nas
```
We will remove those columns

```{r}
data <- select(data,-as.vector(columns_with_too_many_nas$column_name))
```


+ There are many columns with NA values that mostly overlap each other (If there is NA in one of those columns it appears also in the other ones), The following code reflect that fact --> if we filter the NAs from the column: Cross-Street Type --> and we can observe that the number of columns with NAs reduced significantly.


```{r}
filtered_data <- data[!is.na(data$Cross_Street_Type),]


b <- apply(filtered_data, 2, function(c) sum(is.na(c)))
tibble_nas <- tibble(column_name = names(b), number_of_NA = b)
tibble_nas

ggplot(tibble_nas |> filter(number_of_NA >0 ))+
  geom_col(mapping = aes(y=reorder(column_name,number_of_NA), x = number_of_NA)) +
  labs(title = "Number of Nas in the data columns",
       subtitle = "Showing only columns with number of NAs > 0")
```

For now we won't deal with those NAs. Later in the project we would choose the specific columns to analyze together and then we will decide on the best method to deal with the NAs in those columns.


