# Results
<<<<<<< HEAD



=======
```{r}
library(leaflet)
library(leaflet.extras)
library(dplyr)
library(ggplot2)
library(ggmap)
library(lubridate)
```

## Spatial Representation
```{r}
# Sample the data
dt_sampled <- data %>% sample_n(5000)

# Display the random sample
print(dt_sampled)


```

```{r}
# Spatial data


leaflet(dt_sampled) %>%
  addTiles() %>% 
  addHeatmap(
    lat = ~Latitude,
    lng = ~Longitude,
    blur = 15,  # Adjust the blur parameter
    radius = 10)  # Adjust the radius parameter
  
#addMarkers(lng = ~Longitude, lat = ~Latitude, popup = "Marker")

```

## Time Series

```{r}
dt_temporal <- data[,c("Crash_Date/Time","Weather", "Hit/Run", "First_Harmful_Event",
                       "At_Fault")]
dt_temporal <- rename(dt_temporal, "Hit_Run" = `Hit/Run`)

head(dt_temporal)
```


```{r}



#Separate time and Date


dt_temporal$"Crash_Date/Time" <- as.POSIXct(dt_temporal$"Crash_Date/Time", format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")

dt_temporal$Crash_Date <- as.Date(dt_temporal$"Crash_Date/Time", tz = "America/New_York")

#I have a problem with the hours it doesn't show up well
dt_temporal$Crash_Time <- format(dt_temporal$"Crash_Date/Time", "%H:%M:%S")
dt_temporal
#ggplot(dt_temporal, aes(x = as.factor(Crash_week))) +
       #stat_count() +
       #labs(title='Number of car crash per weeks')

```

```{r}
start_date = min(dt_temporal$Crash_Date)
end_date = max(dt_temporal$Crash_Date)

date_sequence <- seq(start_date, end_date, by = "week")
dt_temporal$Crash_week <- cut(dt_temporal$Crash_Date, breaks = c(date_sequence, Inf), labels = as.factor(date_sequence), include.lowest = TRUE)

dt_temporal_red <- dt_temporal %>%
  filter(Crash_Date > as.Date("2013", format = '%Y')) %>%
  group_by(Crash_week) %>%
  mutate(Count_week = n()) %>%
  arrange(Crash_week)

#Create a new table to remove the iteration where there are multiple time the same combination for the columns Crash_week and Count_week.
dt_temporal_week <- dt_temporal_red %>%
  distinct(Crash_week, Count_week, .keep_all = TRUE)%>%
  mutate(Crash_week = as.Date(Crash_week, format = "%Y-%m-%d")) %>%
  mutate(Crash_month = month(Crash_week))
  
dt_temporal_week

ggplot(dt_temporal_week, aes(x = Crash_week, y = Count_week)) +
  geom_line(aes(group=1)) + #because we grouped earlier by Crash_week we need now to specify that we treat all the data as one group.
  geom_point() +
  labs(title='Number of car crash per weeks')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
# Set breaks to 1 month interval
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") 
  #scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")  
```



### Weather Condition

```{r}

dt_temporal <- dt_temporal %>%
  filter(!Weather %in% c("OTHER", "UNKNOWN")) %>%
  filter(!is.na(Weather))

dt_temporal <- dt_temporal %>%
  mutate(Weather = if_else(Weather %in% c("SNOW", "SLEET","BLOWING SNOW"), "SNOW", Weather)) %>%
  mutate(Weather = if_else(Weather %in% c("SEVERE WINDS", "WINTRY MIX", "BLOWING SAND, SOIL, DIRT"), "WIND", Weather))%>%
  mutate(Condition = if_else(Weather %in% c("CLEAR","CLOUDY"),"Good","Bad"))

color_scale <- c("RAINING" = "slateblue", "SNOW" = "whitesmoke", "CLOUDY" = "gray", "FOGGY" ="gray27", "WIND"="seagreen4", "CLEAR" ="skyblue")

dt_temporal$Weather <- reorder(dt_temporal$Weather, X= as.numeric(factor(dt_temporal$Weather,levels = c('CLEAR','CLOUDY','FOGGY','WIND','SNOW','RAINING'))))

ggplot(dt_temporal, aes(x = Condition, fill = Weather)) +
  geom_bar() +
  scale_fill_manual(values = color_scale)

 


#ggplot(dt_temporal, aes(x= Weather))+
#geom_bar()
```
### Hours of the day

```{r}
date <- dt_temporal$Crash_Time[1]
substr(date, )

```


```{r}
# we only care about the hours not the minutes
dt_temporal <- dt_temporal %>%
  mutate(Crash_Time = paste0(substr(Crash_Time, start = 1, stop =2),'h')) 


ggplot(dt_temporal, aes(x=Crash_Time))+
  geom_bar()+
  labs(x = 'Hours in the day',y = 'Number of Crash')
```

### Type of Lane

>>>>>>> 79fe72548c4d5dc8cda3bb0560fedeb39c50b4a9












## Roy
################ Roy

```{r}
data |> head()
```

### Preprocessing for the df_temporal
```{r}
### add some "date slicers" columns
dt_temporal$Week_day <- weekdays(dt_temporal$Crash_Date)
dt_temporal$Month <-  months(dt_temporal$Crash_Date)
dt_temporal$Year <-  year(dt_temporal$Crash_Date)
dt_temporal$first_day_of_month <- as.Date(format(dt_temporal$Crash_Date, "%Y-%m-01"))

library(hms)
dt_temporal$Crash_Time <- as_hms(dt_temporal$Crash_Time)
dt_temporal$Crash_Hour <- hour(dt_temporal$Crash_Time) #Crash hour


### Transform some colkumns to factors
dt_temporal <- mutate(dt_temporal, Week_day = factor(Week_day,
                                             levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

dt_temporal$Month <- factor(dt_temporal$Month, 
                          levels = c("January", "February", "March", "April", "May", "June","July", "August", "September", "October", "November", "December"),)

### arrange the data by date
dt_temporal <- dt_temporal |> arrange(Crash_Date)

dt_temporal_month <- dt_temporal |> group_by(Month, first_day_of_month, Week_day) |> summarise(total_accidents = n())


dt_temporal_Year_month <- dt_temporal |> group_by(Year, Month, first_day_of_month, Week_day) |> summarise(total_accidents = n())


### create a ts object
first_date_in_data <- min(dt_temporal$first_day_of_month)
last_date_in_data <- max(dt_temporal$first_day_of_month)
```


```{r}
ggplot(dt_temporal_month) +
  geom_boxplot(mapping=aes(x=Month, y=total_accidents))



library(ggridges)

### Distribution of total number of accidents by month
ggplot(dt_temporal_month) +
  geom_boxplot(mapping = aes(x = total_accidents, y = Month)) +
  geom_density_ridges(mapping = aes(x = total_accidents, y = Month), alpha = 0.5) +
  theme_minimal() +
  labs(title = "Distribution of total number of accidents by month")


#Distribution of total number of accidents by Day of week
ggplot(dt_temporal_month) +
  geom_boxplot(mapping = aes(x = total_accidents, y = Week_day)) +
  geom_density_ridges(mapping = aes(x = total_accidents, y = Week_day), alpha = 0.5) +
  theme_minimal() +
  labs(title = "Distribution of total number of accidents by weekday")


```

```{r}
# Install and load the necessary packages if not already installed
# install.packages("ggplot2")
# install.packages("ggridges")
# install.packages("patchwork")
library(ggridges)
library(patchwork)


dt_temporal_month <- dt_temporal |> group_by(Month, first_day_of_month, Week_day) |> summarise(total_accidents = n())

dt_temporal_hour <- dt_temporal |> group_by(Month, first_day_of_month, Crash_Hour) |> summarise(total_accidents = n())


# Distribution of total number of accidents by month (the total number of accidents is with the total within in Year-Month combination)
plot_month <- ggplot(dt_temporal_month) +
  geom_boxplot(mapping = aes(x = total_accidents, y = Month)) +
  geom_density_ridges(mapping = aes(x = total_accidents, y = Month), alpha = 0.5) +
  theme_minimal() +
  labs(subtitle = "By month")

# Distribution of total number of accidents by Day of week (the total number of accidents is with the total within in Year-Month combination)
plot_weekday <- ggplot(dt_temporal_month) +
  geom_boxplot(mapping = aes(x = total_accidents, y = Week_day)) +
  geom_density_ridges(mapping = aes(x = total_accidents, y = Week_day), alpha = 0.5) +
  theme_minimal() +
  labs(subtitle = "By weekday")


# Distribution of total number of accidents by Hour (the total number of accidents is with the total within in Year-Month combination)

plot_hours <- ggplot(dt_temporal_hour |> mutate(Crash_Hour = as.factor(Crash_Hour))) +
  geom_boxplot(mapping = aes(x = total_accidents, y = Crash_Hour)) +
  geom_density_ridges(mapping = aes(x = total_accidents, y = Crash_Hour), alpha = 0.5) +
  theme_minimal() +
  labs(subtitle = "By weekday")



# Combine the plots side by side
# combined_plots <- plot_month + plot_weekday + plot_hours

# Combine the plots side by side with one common title
combined_plots <- plot_month + plot_weekday + plot_hours +
  plot_layout(ncol = 2, heights = c(2, 2)) +
  plot_annotation(title = "Distribution of total number of accidents")

# Print the combined plots
print(combined_plots)

```

+ Using the histograms and the density plots together we can get some valuable insights regarding the distribution of total accidents (aggregated by months per yeras) across different months and weekdays.


can see there is an overall different distribution of number of accidents bteween  








```{r}
#dt_temporal_Year_month|>group_by(Year) |> summarise(Total_accidents = sum(total_accidents)

ggplot(dt_temporal_Year_month, aes(x=first_day_of_month, y=total_accidents)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(Week_day~.)



# Filter data for December
december_data <- dt_temporal_Year_month %>%
  filter(month(first_day_of_month) == 12)

highest_points <- dt_temporal_Year_month %>%
  group_by(Year, Week_day) %>%
  slice_max(order_by = total_accidents)

ggplot(dt_temporal_Year_month, aes(x=first_day_of_month, y=total_accidents)) + 
  geom_line() + 
  stat_smooth(method = "loess", span=0.63, col = "blue", size = 1, se = FALSE) +  # Add a moving average line
  geom_point(data = highest_points, col = "black", size = 1.4) +  # Add points only for Jan
  facet_grid(~Week_day) +
  labs(title = "line plot of total accidents faceted by weekday",
       subtitle = "The black dots present for each year the month with \nthe highest number of accidents")


# ggplot(dt_temporal) + 
#   geom_density(stat = "density",
#                aes(x = first_day_of_month), fill = "skyblue", alpha = 0.7) +
#   facet_grid(Week_day~.)




dt_temporal_Year_month
```
In the graph we can see that the nuber of accidents per weekday have a simillar trend, however the magnitude is different - on weekends there are less accidents (reasonable).
It seems that in the middle of he week there are more accidents.
Moreover, we can look at the "black dots" that present for each year the month with the highest number of accident. We can compare those dots between the different week days but also observe the "max-trend" - if we drew another line connecting each one of the black dots, we would get a line trend similar to the loess blue line trend that is presented in the plot.




```{r}
count(data, )
```

- Hit/Run
- NonTraffic
- At_Fault
- Weather
- Surface_Condition
- Light
- Traffic_Control
- Driver_Substance_Abuse
- First_Harmful_Event



```{r}
count(data, Road_Division, Number_of_Lanes)
```


### Hit/run

```{r}
dt_temporal_Year_month_hit_run <- dt_temporal |> group_by(Year, Month, first_day_of_month, Week_day, Hit_Run) |> summarise(total_accidents = n())

dt_temporal_Year_month_hit_run <- dt_temporal |> group_by(Year, Month, first_day_of_month, Week_day, Hit_Run) |> summarise(total_accidents = n())

ggplot(dt_temporal_Year_month_hit_run, aes(x=first_day_of_month, y=total_accidents)) + 
  geom_line() + 
  stat_smooth(method = "loess", span=0.63, col = "blue", size = 1, se = FALSE) +  # Add a moving average line
  facet_grid(~Hit_Run) +
  labs(title = "line plot of total accidents faceted by weekday",
       subtitle = "The black dots present for each year the month with \nthe highest number of accidents")
```



group by --> sum up the number of accident with No and with Yes 
road

##Is it more likely to Hit and Run in a specific Month / Week day / Hour?

```{r}
library(patchwork)

# Filter out NA values for Hit_Run
filtered_data <- dt_temporal %>%
  filter(!is.na(Hit_Run))

# Abbreviate the Month and Weekday names
month_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
weekday_labels <- c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")

# Create a bar plot for Month
plot_month <- ggplot(filtered_data) +
  geom_bar(aes(x = Month, fill = Hit_Run), position = "fill") +
  scale_fill_manual(values = c("No" = "tomato", "Yes" = "darkgreen")) +
  labs(title = "Proportion of Hit-Run Accidents",
       subtitle = "Per Month of Year",
       x = "Month",
       y = "Proportion") +
  scale_x_discrete(labels = month_labels) +  # Abbreviate Month names
  theme_minimal() +
  theme(legend.position = "none")  # Remove the legend

# Create a bar plot for Weekday
plot_weekday <- ggplot(filtered_data) +
  geom_bar(aes(x = Week_day, fill = Hit_Run), position = "fill") +
  scale_fill_manual(values = c("No" = "tomato", "Yes" = "darkgreen")) +
  labs(title = "Proportion of Hit-Run Accidents",
       subtitle = "Per Weekday",
       x = "Weekday",
       y = "Proportion") +
  scale_x_discrete(labels = weekday_labels) +  # Abbreviate Weekday names
  theme_minimal()

# Arrange the two plots side by side
combined_plot <- plot_month + plot_weekday

# Print the combined plot
print(combined_plot)



### Plot Hit-Run by hours
### divide the crash hour into groups
dt_temporal$Crash_Hour_Group <- cut(dt_temporal$Crash_Hour,
  breaks = c(0, 4, 8, 12, 16, 20, 24),
  labels = c("0-4", "4-8", "8-12", "12-16", "16-20", "20-24"),
  include.lowest = TRUE
)


ggplot(dt_temporal %>%
         filter(!is.na(Hit_Run))) +
  geom_bar(aes(x = Crash_Hour_Group, fill = Hit_Run), position = "fill") +
  labs(title = "Proportion of Hit and Run Incidents by Crash Hour",
       x = "Crash Hour",
       y = "Proportion") +
  scale_fill_manual(values = c("No" = "tomato", "Yes" = "darkgreen")) +  # Customize colors if needed
  theme_minimal()

```
+ There doesn't appear to be a correlation between the month and Hit-Run incidents; in other words, there isn't a particular month with a higher likelihood of a Hit-Run accident.
+ However, we observe a higher proportion of Hit-Run incidents during weekends (Saturday and Sunday) compared to the proportions on other weekdays.

+ Concerning the Hit-Run proportion per hour—there is an observed escalation in the proportion of Hit-Run accidents during late hours. Furthermore, there seems to be a consistent "linear trend" in Hit-Run incidents as the day unfolds.

```{r}
ggplot()

+ geom_bar(position = "dodge")
```



### First_Harmful_Event

```{r}
#count(dt_temporal, First_Harmful_Event) |> arrange(desc(n))

### taking only the 7 top categories in  First_Harmful_Event
df_for_harmful_event <- dt_temporal %>%
  filter(!is.na(First_Harmful_Event)) %>%  # Remove NA values
  mutate(
    First_Harmful_Event = tolower(First_Harmful_Event),
    First_Harmful_Event = factor(First_Harmful_Event) %>% fct_lump(n = 7, other_level = "Other"))

###Mosaic plot 1:
mosaicplot(table(factor(df_for_harmful_event$First_Harmful_Event), df_for_harmful_event$Hit_Run), main= "Mosaic plot for Hit Run vs First Harmful Event",
           color = c('tomato', 'darkgreen'), 
           las = 2  # Rotate labels by 90 degrees
)

count(df_for_harmful_event, Hit_Run, First_Harmful_Event)


#+
 # scale_fill_manual(values = c("No" = "tomato", "Yes" = "darkgreen")) +

```
We can see that the first harmful event has a great connection / correlation to the Hit-Run variable --> there are great variations between the proportion of Hit-Run vs. Not Hit-Run accidents given the different `First_Harmful_Event` categories.

+ Note that for "Parked-Vehicle" the proportion of Hit-Run is 0.45 which is more than twice as big as the overall Hit-Run proportion (0.18):

```{r}
dt_temporal |> count(Hit_Run) |> mutate("prop" = n/nrow(dt_temporal))
a <- dt_temporal |> filter(grepl(c("Parked"), First_Harmful_Event, ignore.case = T)) |> count(Hit_Run, First_Harmful_Event) 
a|> mutate("prop" = n/sum(a$n)) 

```


Lets take the biggest category: other_vehicle and analyze it even further

```{r}

df_mosaic_by_hour_harmful <- df_for_harmful_event |> filter(grepl(c("Parked"), First_Harmful_Event, ignore.case = T)) |> group_by(First_Harmful_Event, Crash_Hour, Hit_Run) |> summarise(Freq = n())


vcd:: mosaic(Hit_Run ~ First_Harmful_Event + Crash_Hour,
 data = df_mosaic_by_hour_harmful |> 
   mutate(
          First_Harmful_Event = as.character(First_Harmful_Event)),
 direction = c("v", "v", "h"), 
 highlighting_fill=c('tomato','darkgreen'))

```




## At_Fault

```{r}
dt_temporal

df_mosaic_by_weather <- dt_temporal |> group_by(Weather, Hit_Run) |> summarise(Freq = n())


###Mosaic plot 1:
mosaicplot(table(dt_temporal$Weather, dt_temporal$Hit_Run), main= "Mosaic plot for Hit Run vs First Harmful Event",
           color = c('tomato', 'darkgreen'), 
           las = 2  # Rotate labels by 90 degrees
)


Weather

vcd:: mosaic(Hit_Run ~ Weather,
 data = df_mosaic_by_weather,
 direction = c("v", "h"), 
 highlighting_fill=c('tomato','darkgreen'))
```


## Things we found from graphs we created"
+ We checked and there doesn't seem to be a correlation between the `Weather` and the `Hit_Run`




```{r}
dt_temporal_week


ts_data <- ts(data = dt_temporal_week$Crash_week, start = c(2015,1,3), end = c(2023,11,13), frequency = 51)

#dt_temporal_week$Crash_Date[nrow(dt_temporal_week)]



ts_data<- ts(df_temporal_week$Crash_week)
plot(decompose())


```



Box plot 
x axis - day of week
the box itself will use the number of accidents









