# Results
<<<<<<< HEAD



=======
```{r}
library(leaflet)
library(leaflet.extras)
library(dplyr)
library(ggplot2)
library(ggmap)
library(lubridate)
```

## Spatial Representation
```{r}
# Sample the data
dt_sampled <- data %>% sample_n(5000)

# Display the random sample
print(dt_sampled)


```

```{r}
# Spatial data


leaflet(dt_sampled) %>%
  addTiles() %>% 
  addHeatmap(
    lat = ~Latitude,
    lng = ~Longitude,
    blur = 15,  # Adjust the blur parameter
    radius = 10)  # Adjust the radius parameter
  
#addMarkers(lng = ~Longitude, lat = ~Latitude, popup = "Marker")

```

## Time Series

```{r}
dt_temporal <- data[,c("Crash_Date/Time","Weather")]
head(dt_temporal)
```


```{r}



#Separate time and Date


dt_temporal$"Crash_Date/Time" <- as.POSIXct(dt_temporal$"Crash_Date/Time", format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")

dt_temporal$Crash_Date <- as.Date(dt_temporal$"Crash_Date/Time", tz = "America/New_York")

#I have a problem with the hours it doesn't show up well
dt_temporal$Crash_Time <- format(dt_temporal$"Crash_Date/Time", "%H:%M:%S")
dt_temporal
#ggplot(dt_temporal, aes(x = as.factor(Crash_week))) +
       #stat_count() +
       #labs(title='Number of car crash per weeks')

```

```{r}
start_date = min(dt_temporal$Crash_Date)
end_date = max(dt_temporal$Crash_Date)

date_sequence <- seq(start_date, end_date, by = "week")
dt_temporal$Crash_week <- cut(dt_temporal$Crash_Date, breaks = c(date_sequence, Inf), labels = as.factor(date_sequence), include.lowest = TRUE)

dt_temporal_red <- dt_temporal %>%
  filter(Crash_Date > as.Date("2013", format = '%Y')) %>%
  group_by(Crash_week) %>%
  mutate(Count_week = n()) %>%
  arrange(Crash_week)

#Create a new table to remove the iteration where there are multiple time the same combination for the columns Crash_week and Count_week.
dt_temporal_week <- dt_temporal_red %>%
  distinct(Crash_week, Count_week, .keep_all = TRUE)%>%
  mutate(Crash_week = as.Date(Crash_week, format = "%Y-%m-%d")) %>%
  mutate(Crash_month = month(Crash_week))
  
dt_temporal_week

ggplot(dt_temporal_week, aes(x = Crash_week, y = Count_week)) +
  geom_line(aes(group=1)) + #because we grouped earlier by Crash_week we need now to specify that we treat all the data as one group.
  geom_point() +
  labs(title='Number of car crash per weeks')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
# Set breaks to 1 month interval
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") 
  #scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")  
```


```{r}
dt_temporal_week

ts_data<- ts(df_temporal_week$Crash_week)
plot(decompose())


```

### Weather Condition

```{r}

dt_temporal <- dt_temporal %>%
  filter(!Weather %in% c("OTHER", "UNKNOWN")) %>%
  filter(!is.na(Weather))

dt_temporal <- dt_temporal %>%
  mutate(Weather = if_else(Weather %in% c("SNOW", "SLEET","BLOWING SNOW"), "SNOW", Weather)) %>%
  mutate(Weather = if_else(Weather %in% c("SEVERE WINDS", "WINTRY MIX", "BLOWING SAND, SOIL, DIRT"), "WIND", Weather))%>%
  mutate(Condition = if_else(Weather %in% c("CLEAR","CLOUDY"),"Good","Bad"))

color_scale <- c("RAINING" = "slateblue", "SNOW" = "whitesmoke", "CLOUDY" = "gray", "FOGGY" ="gray27", "WIND"="seagreen4", "CLEAR" ="skyblue")

dt_temporal$Weather <- reorder(dt_temporal$Weather, X= as.numeric(factor(dt_temporal$Weather,levels = c('CLEAR','CLOUDY','FOGGY','WIND','SNOW','RAINING'))))

ggplot(dt_temporal, aes(x = Condition, fill = Weather)) +
  geom_bar() +
  scale_fill_manual(values = color_scale)

 


#ggplot(dt_temporal, aes(x= Weather))+
#geom_bar()
```
### Hours of the day

```{r}
date <- dt_temporal$Crash_Time[1]
substr(date, )

```


```{r}
# we only care about the hours not the minutes
dt_temporal <- dt_temporal %>%
  mutate(Crash_Time = paste0(substr(Crash_Time, start = 1, stop =2),'h')) 


ggplot(dt_temporal, aes(x=Crash_Time))+
  geom_bar()+
  labs(x = 'Hours in the day',y = 'Number of Crash')
```

### Type of Lane

>>>>>>> 79fe72548c4d5dc8cda3bb0560fedeb39c50b4a9
